---
title: "Tools to work fast with NetCDF"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{NetCDF tools}
---

The [**rbeni**](https://github.com/stineb/rbeni) package contains a set of functions that let you do standard tasks working with NetCDF files on one line and let you quickly convert the contents of a NetCDF file into a tidy data frame (tibble).

Here is a demonstration example.

## Read

Load the package and read the contents of one NetCDF file.
```{r message=FALSE}
library(rbeni)
library(tidyverse)  # depends
nc <- read_nc_onefile("~/data/stocker20gmd_outputs/global/global_FULL_fAPAR3g_v2_2000_2016.a.gpp.nc")
```

The returned object `nc` is a list with:
```{r}
print(ls(nc))
```

- `lon`: vector of longitudes of gridcell mid-points
- `lat`: vector of latitudes of gridcell mid-points
- `time`: vector of dates created by `lubridate::ymd dates`
- `varnams`: vector of all variable names as strings, 
- `vars`: named list of the data arrays (lon x lat [x time]) for each variable. The names of that list correspond to `varnams`.

```{r}
## available variable names
print(nc$varnams)

## available variable arrays (names thereof)
print(ls(nc$vars))

## dimension of variable array number one
print(dim(nc$vars[[1]]))
```

## Make tidy

The object `nc` can be converted into a tidy dataframe.
```{r eval=FALSE}
df <- nc_to_df(nc, varnam = "gpp")
head(df)
```
The data frame is organised by gridcells along rows. Since the NetCDF contains multiple time steps, the time series are nested inside the column `data` for each gridcell.
```{r eval=FALSE}
head(df$data[[1]])
```
Note that `df` contains rows for all gridcells, also those that have no data (here, non-land gridcells). That's why the above time series data frame is empty.

Since object `nc` contains data for multiple time steps, we can select a single time step and get an object that has only lon and lat.
```{r}
nc_2d <- slice_nc(nc, 1)
```

This 2D data frame can be converted back into a grid:
```{r eval=FALSE}
gridded <- df_to_grid(nc_2d, varnam = "data")
gridded
```

(This makes sense only if `df` is not a nested data frame and contains only data for one time step.)

## Plot

This can now be used to create a nice map.
```{r message=FALSE}
gg <- plot_map3(nc_2d, varnam = "gpp", plot_title = "GPP", latmin = -65, latmax = 85)
gg
```

## Write

Very useful also to quickly write the `nc` object into a new NetCDF file.
```{r}
write_nc2(nc, path = "./test.nc")
```
